acquisition of TEs

### 2021-06-15 simulations_old_versions/Analyze_TEs.Rmd
organizing folders

library(corrplot)
corrplot(mycor)
write.csv(mycor, "Output/ObservedTE_correlations.csv", col.names=NA)

cor.test(Correlation$TE, Correlation$B, method = "pearson")
cor.test(Correlation$TE, Correlation$W, method = "pearson")
cor.test(Correlation$TE, Correlation$D, method = "pearson")
cor.test(Correlation$TE, Correlation$N, method = "pearson")
cor.test(Correlation$TE, Correlation$V, method = "pearson")
cor.test(Correlation$DI, Correlation$B, method = "pearson")
cor.test(Correlation$NI, Correlation$B, method = "pearson")

@@

## clean the simulation data
data_simulation1 <- data_simulation1 %>%
  arrange(DOM, NONDOM) %>%
  filter(NONDOM <= DOM) %>%
  # Get word vocabulary W
  mutate(WORD = DOM + NONDOM) %>%
  # Create other variables
  mutate(BALANCE = NONDOM/(DOM+NONDOM), 
         TE = (DOM*NONDOM)/LEARNABLE) %>% # If DOM and NONDOM are independent, TE = DOM*NONDOM/LEARNABLE
  # Round to keep whole numbers only
  mutate_at(vars(-BALANCE), round, 0) %>%
  # bin B into 6 groups
  mutate(bin_BALANCE = cut(BALANCE, breaks = 5, 
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  arrange(DOM, NONDOM, bin_BALANCE)

## visualizing Simulation 1
simulation1 <- data_simulation1 %>%
  # convert to long data
  pivot_longer(c(WORD, DOM, NONDOM), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("DOM", "NONDOM", "WORD"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             DOM = "Panel 1A: Dominant vocabulary \n (DOM)", 
                             NONDOM = "Panel 1B: Non-dominant vocabulary \n (NONDOM)",
                             WORD = "Panel 1C: Total vocabulary \n (WORD)")) %>%
  ggplot(aes(x = number, y = TE, color = bin_BALANCE, linetype = bin_BALANCE)) +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) +
  #scale_color_manual(values=c("#5e4fa2", "#3288bd", "#abdda4", "#fee08b", "#fc8d59", "#d53e4f")) + 
  #scale_color_discrete(direction = -1) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  ylim(0, 600) +
  xlim(0, 1000) +
  labs(x = "Number of words", 
       y = "Total number of translation equivalents (TE)", 
       color = "Vocabulary balance (BALANCE)",
       linetype = "Vocabulary balance (BALANCE)",
       title = "1) Simulated data") +
  guides(color = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

@@

## Observed data: Testing Simulation 1
```{r}
Observed_plot_inputs_TE <- keepers_ws_TE %>% 
  select(baby_id, ID_testdate, age_days, number_of_te, word_vocab, TotalNW_Dom, TotalNW_NonDom, balance_vocab) %>%
  pivot_longer(c(word_vocab, TotalNW_Dom, TotalNW_NonDom), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("TotalNW_Dom", "TotalNW_NonDom", "word_vocab"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             TotalNW_Dom = "Panel 2A: Dominant vocabulary \n (DOM)", 
                             TotalNW_NonDom = "Panel 2B: Non-dominant vocabulary \n (NONDOM)",
                             word_vocab = "Panel 2C: Total vocabulary \n (WORD)")) %>%
  mutate(bin_BALANCE = cut(balance_vocab, breaks = 5, 
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  #mutate(bin_B = case_when(balance_vocab < 0.2 ~ "0.1",
  #                         balance_vocab >= 0.2 & balance_vocab < 0.3 ~ "0.2",
  #                         balance_vocab >= 0.3 & balance_vocab < 0.4 ~ "0.3",
  #                         balance_vocab >= 0.4 & balance_vocab < 0.5 ~ "0.5",
  #                         TRUE ~ "0.5")) %>%
  ggplot(aes(x = number, y = number_of_te, color = bin_BALANCE, linetype = bin_BALANCE)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        legend.position = "bottom",
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  ylim(0, 600) +
  xlim(0, 1000) +
  labs(x = "Number of words", 
       y = "Total number of translation equivalents (TE)", 
       color = "Vocabulary balance (BALANCE)",
       linetype = "Vocabulary balance (BALANCE)",
       title = "2) Observed data") +
  guides(colour = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

@@

### one-way ANOVA to compare the 5 balance groups
aov_TE_binBALANCE <- keepers_ws_TE %>% 
  mutate(bin_BALANCE = cut(balance_vocab, breaks = 5, 
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  aov(number_of_te ~ bin_BALANCE, data = .)

@@

check <- keepers_ws_TE %>% 
  mutate(bin_BALANCE = cut(balance_vocab, breaks = 5, 
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5")))

pairwise.t.test(check$number_of_te, check$bin_BALANCE)

mean_TE_binBALANCE <- keepers_ws_TE %>% 
  mutate(bin_BALANCE = cut(balance_vocab, breaks = 5, 
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  group_by(bin_BALANCE) %>%
  summarize(n = n(),
            mean_TE = mean(number_of_te, na.rm = T),
            sd_TE = sd(number_of_te, na.rm = T),
            min_TE = min(number_of_te, na.rm = T),
            max_TE = max(number_of_te, na.rm = T),)

@@

# One-way Anova to compare the 3 V groups
keepers_ws_TE_3Vsubset <- keepers_ws_TE %>%
  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months (V = 244.9 - 451.9)",
                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months (V = 491.8 - 604.1)",
                              TRUE ~ "28-33 months (V = 620.4 - 638.9)"))

summary(aov(concept_vocab ~ V_subset, data = keepers_ws_TE_3Vsubset))

mean_conceptV_Vsubset <- keepers_ws_TE_3Vsubset %>%
  group_by(V_subset) %>%
  summarise(mean_conceptV = mean(concept_vocab, na.rm = T),
            sd_conceptV = sd(concept_vocab, na.rm = T),
            min_conceptV = min(concept_vocab, na.rm = T),
            max_conceptV = max(concept_vocab, na.rm = T))

pairwise.t.test(keepers_ws_TE_3Vsubset$concept_vocab, keepers_ws_TE_3Vsubset$V_subset,
                 p.adjust.method = "BH")

## Mean proportion of TE
stacked_byV_wide <- keepers_ws_TE %>%
  select(baby_id, age_months_percentile, balance_vocab, word_vocab, concept_vocab, 
         number_of_te, Dom_unique_words, NonDom_unique_words) %>%
  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months",
                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months",
                              TRUE ~ "28-33 months")) %>%
  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "Most Balanced",
                               balance_vocab <= 0.20 ~ "Least Balanced",
                               TRUE ~ "Medium Balanced"))

stacked_byV_wide %>%
  group_by(V_subset) %>%
  summarise(mean_propTE = mean(propTE, na.rm=T), 
            sd_propTE = sd(propTE, na.rm=T),
            min_propTE = min(propTE, na.rm=T),
            max_propTE = max(propTE, na.rm=T))

# Anova to compare the number of TEs among the 3 developmental groups in the stacked barchart
stacked_byV_wide %>%  
  group_by(V_subset) %>%
  summarise(mean_TE = mean(number_of_te, na.rm=T), 
            sd_TE = sd(number_of_te, na.rm=T),
            min_TE = min(number_of_te, na.rm=T),
            max_TE = max(number_of_te, na.rm=T))

TE_byV_anova <- aov(number_of_te ~ V_subset, data = stacked_byV_wide)
summary(TE_byV_anova)

pairwise.t.test(stacked_byV_wide$number_of_te, stacked_byV_wide$V_subset,
                 p.adjust.method = "BH")

## Anova to compare the unique dom vocab size and unique non-dom vocab size among the 3 developmental groups in the stacked barchart
stacked_byV_wide %>%  
  group_by(BalSubset) %>%
  summarise(mean_DI = mean(Dom_unique_words, na.rm=T), 
            sd_DI = sd(Dom_unique_words, na.rm=T),
            min_DI = min(Dom_unique_words, na.rm=T),
            max_DI = max(Dom_unique_words, na.rm=T))

DI_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_byV_wide)
summary(DI_byB_anova)

pairwise.t.test(stacked_byV_wide$Dom_unique_words, stacked_byV_wide$BalSubset,
                 p.adjust.method = "BH")

stacked_byV_wide %>% 
  group_by(BalSubset) %>%
  summarise(mean_NI = mean(NonDom_unique_words, na.rm=T), 
            sd_NI = sd(NonDom_unique_words, na.rm=T),
            min_NI = min(NonDom_unique_words, na.rm=T),
            max_NI = max(NonDom_unique_words, na.rm=T))

NI_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_byV_wide)
summary(NI_byB_anova)

pairwise.t.test(stacked_byV_wide$NonDom_unique_words, stacked_byV_wide$BalSubset,
                 p.adjust.method = "BH")

@@

data_simulation3 <- do.call(rbind, lapply(LEARNABLE, simulation3)) %>%
  arrange(LEARNABLE, DOM, NONDOM) %>%
  filter(NONDOM <= DOM) %>% # eliminate cases when NONDOM > DOM
  mutate(WORD = DOM+NONDOM, # word vocabulary
         BALANCE = NONDOM/(DOM+NONDOM), # balance 
         TE = DOM*NONDOM/LEARNABLE) %>% # If DOM and NONDOM are independent, TE = DOM*NONDOM/LEARNABLE
  mutate_at(vars(-BALANCE), round, 0) %>%
  mutate(bin_BALANCE = cut(BALANCE, breaks = 5, 
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  arrange(bin_BALANCE)

@@

# Testing Simulation 3: Figures 7 & 8
## Theoretical distribution of TEs (Simulation): D * N / V (TotalNW_Dom * TotalNW_NonDom / Average_90percentile) 
```{r}
## Create predicted TEs for simulation
keepers_ws_TE_simulation <- keepers_ws_TE %>%
  mutate(Predicted_TE_percentile = TotalNW_Dom * TotalNW_NonDom/Average_90percentile) %>%
  select(baby_id, ID_testdate, Predicted_TE_percentile, number_of_te, Average_90percentile, TotalNW_Dom, TotalNW_NonDom, word_vocab, age_months_binned)

## LM model
Percentile_model <- keepers_ws_TE_simulation %>%
  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)

summary(Percentile_model)

## Ploting Y = observed rawTE  & X = Predicted rawTE 
TEprediction_percentile <- keepers_ws_TE_simulation %>%
  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
#  facet_grid(. ~TEgroup) +
  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
#  geom_text(aes(label=ID_testdate)) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,450) +
  ylim(-10,450) +
  theme_minimal()

ggsave(here("Output/TEprediction_percentile.png"), TEprediction_percentile,
       height=5, width=5)

## Identifying outliers
CheckOutliers_model <- keepers_ws_TE_simulation %>%
  lm(number_of_te ~ 0 + Predicted_TE_percentile, 
#  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom,
     data = .)
summary(CheckOutliers_model)

@@

## Ploting Y = observed rawTE  & X = Predicted rawTE 
TEprediction_percentile_NoOutliers <- keepers_ws_TE_simulation_removeCD %>%
  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
#  facet_grid(. ~TEgroup) +
  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,430) +
  ylim(-10,430) +
  theme_minimal()

Percentile_model_NoOutliers_LessThan300W <- keepers_ws_TE_simulation_removeCD %>%
  filter(word_vocab < 300) %>%
  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)

summary(Percentile_model_NoOutliers_LessThan300W)

Percentile_model_NoOutliers_MoreThan300W <- keepers_ws_TE_simulation_removeCD %>%
  filter(word_vocab > 300) %>%
  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)

summary(Percentile_model_NoOutliers_MoreThan300W)

## Looking closer at the infants with total vocabulary < 300

keepers_ws_TE_simulation_LessThan300 <- keepers_ws_TE_simulation %>%
  filter(word_vocab < 300)

stats::ks.test(keepers_ws_TE_simulation_LessThan300$number_of_te,
               keepers_ws_TE_simulation_LessThan300$Predicted_TE_percentile)

keepers_ws_TE_simulation_MoreThan300 <- keepers_ws_TE_simulation %>%
  filter(word_vocab > 300)

stats::ks.test(keepers_ws_TE_simulation_MoreThan300$number_of_te,
               keepers_ws_TE_simulation_MoreThan300$Predicted_TE_percentile)

# Plot
TEprediction_percentile_facet_lessW300 <- keepers_ws_TE_simulation %>%
  mutate(W_group = if_else(word_vocab < 300, "Less than 300 total vocabulary", "More than 300 total vocabulary")) %>%
  filter(W_group == "Less than 300 total vocabulary") %>%
  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs",
       title = "Less than 300 total vocabulary") +
  xlim(0,80) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

@@

ggsave(here("Output/TEprediction_percentile_facetW.png"), TEprediction_percentile_facetByW300,
       height=5,width=10)

Percentile_model_LessThan300W <- keepers_ws_TE_simulation %>%
  filter(word_vocab < 300) %>%
  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)

summary(Percentile_model_LessThan300W)

Percentile_model_MoreThan300W <- keepers_ws_TE_simulation %>%
  filter(word_vocab > 300) %>%
  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)

summary(Percentile_model_MoreThan300W)

>>>>

library(corrplot)
corrplot(mycor)
write.csv(mycor, "Output/ObservedTE_correlations.csv", col.names=NA)
```

@@

## clean the simulation data
data_simulation1 <- data_simulation1 %>%
  arrange(DOM, NONDOM) %>%
  filter(NONDOM <= DOM) %>%
  # Get word vocabulary W
  mutate(WORD = DOM + NONDOM) %>%
  # Create other variables
  mutate(BALANCE = NONDOM/(DOM+NONDOM), 
         TE = (DOM*NONDOM)/LEARNABLE) %>% # If DOM and NONDOM are independent, TE = DOM*NONDOM/LEARNABLE
  # Round to keep whole numbers only
  mutate_at(vars(-BALANCE), round, 0) %>%
  # bin B into 5 groups
  mutate(bin_BALANCE = cut(BALANCE, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  arrange(DOM, NONDOM, bin_BALANCE)

## visualizing Simulation 1
simulation1 <- data_simulation1 %>%
  # convert to long data
  pivot_longer(c(WORD, DOM, NONDOM), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("DOM", "NONDOM", "WORD"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             DOM = "Panel 1A: Dominant vocabulary \n (DOM)", 
                             NONDOM = "Panel 1B: Non-dominant vocabulary \n (NONDOM)",
                             WORD = "Panel 1C: Total vocabulary \n (WORD)")) %>%
  ggplot(aes(x = number, y = TE, color = bin_BALANCE, linetype = bin_BALANCE)) +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  ylim(0, 600) +
  xlim(0, 1000) +
  labs(x = "Number of words", 
       y = "Total number of translation equivalents (TE)", 
       color = "Vocabulary balance (BALANCE)",
       linetype = "Vocabulary balance (BALANCE)",
       title = "1) Simulated data") +
  guides(color = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

@@

## Observed data: Testing Simulation 1
```{r}
Observed_plot_inputs_TE <- keepers_ws_TE %>% 
  select(baby_id, ID_testdate, age_days, number_of_te, word_vocab, TotalNW_Dom, TotalNW_NonDom, balance_vocab) %>%
  pivot_longer(c(word_vocab, TotalNW_Dom, TotalNW_NonDom), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("TotalNW_Dom", "TotalNW_NonDom", "word_vocab"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             TotalNW_Dom = "Panel 2A: Dominant vocabulary \n (DOM)", 
                             TotalNW_NonDom = "Panel 2B: Non-dominant vocabulary \n (NONDOM)",
                             word_vocab = "Panel 2C: Total vocabulary \n (WORD)")) %>%
  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) 
  ggplot(aes(x = number, y = number_of_te, color = bin_BALANCE, linetype = bin_BALANCE)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        legend.position = "bottom",
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  ylim(0, 600) +
  xlim(0, 1000) +
  labs(x = "Number of words", 
       y = "Total number of translation equivalents (TE)", 
       color = "Vocabulary balance (BALANCE)",
       linetype = "Vocabulary balance (BALANCE)",
       title = "2) Observed data") +
  guides(colour = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

@@

# testing Prediction 1a: To compare TE among the 5 BALANCE groups 
## correlation between TE and BALANCE
cor.test(Correlation$TE, Correlation$BALANCE, method = "pearson")

## one-way ANOVA
aov_TE_binBALANCE <- keepers_ws_TE %>% 
  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  aov(number_of_te ~ bin_BALANCE, data = .)

@@

check <- keepers_ws_TE %>% 
  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5")))

pairwise.t.test(check$number_of_te, check$bin_BALANCE,
                p.adjust.method = "BH")

mean_TE_binBALANCE <- keepers_ws_TE %>% 
  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  group_by(bin_BALANCE) %>%
  summarise(n = n(),
            mean_TE = mean(number_of_te, na.rm = T),
            sd_TE = sd(number_of_te, na.rm = T),
            min_TE = min(number_of_te, na.rm = T),
            max_TE = max(number_of_te, na.rm = T)) %>%
  mutate(across(!bin_BALANCE & !sd_TE, ~ round(.,0))) %>%
  mutate_at(vars(sd_TE), ~ round(.,2))

# testing Prediction 1b: To compare TE and WORD or DOM
cor.test(Correlation$TE, Correlation$WORD, method = "pearson")
cor.test(Correlation$TE, Correlation$DOM, method = "pearson")

# testing Prediction 1c: To compare TE and NONDOM
cor.test(Correlation$TE, Correlation$NONDOM, method = "pearson")
```

@@

# testing Prediction 2a: To compare CONCEPT among the 3 LEARNABLE groups in the stacked bar chart
## correlation between age (in months) and CONCEPT
cor.test(Correlation$CONCEPT, Correlation$Age_month, method = "pearson")

## one-way ANOVE to compare CONCEPT among LEARNABLE
keepers_ws_TE_3LEARNABLEsubset <- keepers_ws_TE %>%
  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months (V = 244.9 - 451.9)",
                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months (V = 491.8 - 604.1)",
                              TRUE ~ "28-33 months (V = 620.4 - 638.9)"))

summary(aov(concept_vocab ~ LEARNABLE_subset, data = keepers_ws_TE_3LEARNABLEsubset))

mean_concept_LEARNABLEsubset <- keepers_ws_TE_3LEARNABLEsubset %>%
  group_by(LEARNABLE_subset) %>%
  summarise(mean_concept = mean(concept_vocab, na.rm = T),
            sd_concept = sd(concept_vocab, na.rm = T),
            min_concept = min(concept_vocab, na.rm = T),
            max_concept = max(concept_vocab, na.rm = T)) %>%
  mutate(across(is.numeric, ~ round(.,1)))

pairwise.t.test(keepers_ws_TE_3LEARNABLEsubset$concept_vocab, keepers_ws_TE_3LEARNABLEsubset$LEARNABLE_subset,
                p.adjust.method = "BH")

# testing Prediction 2b: To compare TE among the 3 LEARNABLE groups in the stacked bar chart
## one-way ANOVA to compare TE among LEARNABLE
stacked_byLEARNABLE_wide <- keepers_ws_TE %>%
  select(baby_id, age_months_percentile, balance_vocab, word_vocab, concept_vocab, 
         number_of_te, Dom_unique_words, NonDom_unique_words) %>%
  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months",
                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months",
                              TRUE ~ "28-33 months")) %>%
  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "Most Balanced",
                               balance_vocab <= 0.20 ~ "Least Balanced",
                               TRUE ~ "Medium Balanced"))

TE_byLEARNABLE_anova <- aov(number_of_te ~ LEARNABLE_subset, data = stacked_byLEARNABLE_wide)
summary(TE_byLEARNABLE_anova)

pairwise.t.test(stacked_byLEARNABLE_wide$number_of_te, stacked_byLEARNABLE_wide$LEARNABLE_subset,
                 p.adjust.method = "BH")

mean_TE_LEARNABLEsubset <- stacked_byLEARNABLE_wide %>%  
  group_by(LEARNABLE_subset) %>%
  summarise(mean_TE = mean(number_of_te, na.rm=T), 
            sd_TE = sd(number_of_te, na.rm=T),
            min_TE = min(number_of_te, na.rm=T),
            max_TE = max(number_of_te, na.rm=T)) %>%
  mutate(across(is.numeric, ~ round(.,1)))

## correlation between age (in months) and TE
cor.test(Correlation$TE, Correlation$Age_month, method = "pearson")

# testing Prediction 2c: One-way Anova to compare DOM-SINGLET among the 3 LEARNABLE groups in the stacked bar chart
## correlation between DOM-SINGLET and age (in month)
cor.test(Correlation$"DOM-SINGLET", Correlation$Age_month, method = "pearson")

## correlation between DOM-SINGLET and BALANCE
cor.test(Correlation$"DOM-SINGLET", Correlation$BALANCE, method = "pearson")

## one-way ANOVA to compare DOM-SINGLET among LEARNABLE
DOMSINGLET_byBALANCE_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_byLEARNABLE_wide)
summary(DOMSINGLET_byBALANCE_anova)

pairwise.t.test(stacked_byLEARNABLE_wide$Dom_unique_words, stacked_byLEARNABLE_wide$BalSubset,
                 p.adjust.method = "BH")

mean_DOMSINGLET_LEARNABLEsubset <- stacked_byLEARNABLE_wide %>%  
  group_by(BalSubset) %>%
  summarise(mean_DOM_SINGLET = mean(Dom_unique_words, na.rm=T), 
            sd_DOM_SINGLET = sd(Dom_unique_words, na.rm=T),
            min_DOM_SINGLET = min(Dom_unique_words, na.rm=T),
            max_DOM_SINGLET = max(Dom_unique_words, na.rm=T)) %>%
  mutate(across(is.numeric, ~ round(.,1)))

# testing Prediction 2d: One-way Anova to compare NONDOM-SINGLET among the 3 LEARNABLE groups in the stacked barchart
## correlation between NONDOM-SINGLET and age (in month)
cor.test(Correlation$"NONDOM-SINGLET", Correlation$Age_month, method = "pearson")

## correlation between NONDOM-SINGLET and BALANCE
cor.test(Correlation$"NONDOM-SINGLET", Correlation$BALANCE, method = "pearson")

## one-way ANOVA to compare NONDOM-SINGLET among LEARNABLE
NONDOMSINGLET_byBALANCE_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_byLEARNABLE_wide)
summary(NONDOMSINGLET_byBALANCE_anova)

pairwise.t.test(stacked_byLEARNABLE_wide$NonDom_unique_words, stacked_byLEARNABLE_wide$BalSubset,
                 p.adjust.method = "BH")

mean_NONDOMSINGLET_LEARNABLEsubset <-stacked_byLEARNABLE_wide %>% 
  group_by(BalSubset) %>%
  summarise(mean_NONDOM_SINGLET = mean(NonDom_unique_words, na.rm=T), 
            sd_NONDOM_SINGLET = sd(NonDom_unique_words, na.rm=T),
            min_NONDOM_SINGLET = min(NonDom_unique_words, na.rm=T),
            max_NONDOM_SINGLET = max(NonDom_unique_words, na.rm=T)) %>%
  mutate(across(is.numeric, ~ round(.,1)))
```

@@

data_simulation3 <- do.call(rbind, lapply(LEARNABLE, simulation3)) %>%
  arrange(LEARNABLE, DOM, NONDOM) %>%
  filter(NONDOM <= DOM) %>% # eliminate cases when NONDOM > DOM
  mutate(WORD = DOM+NONDOM, # word vocabulary
         BALANCE = NONDOM/(DOM+NONDOM), # balance 
         TE = DOM*NONDOM/LEARNABLE) %>% # If DOM and NONDOM are independent, TE = DOM*NONDOM/LEARNABLE
  mutate_at(vars(-BALANCE), round, 0) %>%
  mutate(bin_BALANCE = cut(BALANCE, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  arrange(bin_BALANCE)

@@

# Testing Simulation 3: Figures 7 & 8
## Theoretical distribution of TEs (Simulation): D * N / V (TotalNW_Dom * TotalNW_NonDom / Average_90percentile) 
```{r}
# testing Prediction 3: linear regression model comparing simulated and observed TE
## creating predicted TEs for simulation
keepers_ws_TE_simulation <- keepers_ws_TE %>%
  mutate(Predicted_TE_percentile = TotalNW_Dom * TotalNW_NonDom/Average_90percentile) %>%
  select(baby_id, ID_testdate, Predicted_TE_percentile, number_of_te, Average_90percentile, TotalNW_Dom, TotalNW_NonDom, word_vocab, age_months_binned)

##linear regression model
Percentile_model <- keepers_ws_TE_simulation %>%
  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)
summary(Percentile_model)

## ploting Y = observed rawTE  & X = Predicted rawTE 
plot_TEprediction_percentile <- keepers_ws_TE_simulation %>%
  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,450) +
  ylim(-10,450) +
  theme_minimal()

ggsave(here("Output/plot_Simulation3_data.png"), plot_TEprediction_percentile,
       height=5, width=5)

## identifying outliers
CheckOutliers_model <- keepers_ws_TE_simulation %>%
  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom,
     data = .)
summary(CheckOutliers_model)

@@

# splitting the model for children with WORD < 300 and those with WORD > 300
## linear regression model for those with WORD > 300
Percentile_model_MoreThan300W <- keepers_ws_TE_simulation %>%
  filter(word_vocab > 300) %>%
  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)
summary(Percentile_model_NoOutliers_MoreThan300W)

## linear regression model for those with WORD < 300
Percentile_model_LessThan300W <- keepers_ws_TE_simulation %>%
  filter(word_vocab < 300) %>%
  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)
summary(Percentile_model_NoOutliers_LessThan300W)

## plotting the two separate models (those with WORD < 300 + those with WORD > 300)
TEprediction_percentile_facet_lessW300 <- keepers_ws_TE_simulation %>%
  mutate(W_group = if_else(word_vocab < 300, "Less than 300 total vocabulary", "More than 300 total vocabulary")) %>%
  filter(W_group == "Less than 300 total vocabulary") %>%
  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs",
       title = "Less than 300 total vocabulary") +
  xlim(0,80) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

@@

ggsave(here("Output/plot_Simulation3_data_2WORDgroups.png"), TEprediction_percentile_facetByW300,
       height=5,width=10)
```
